### 2023-06-29 journal/conclusion.tex
Updates to related work--incorporating new info about Jayanti et al.'s lower bound

% !TEX root =  podc-submission.tex

@@

A gap remains between our queue, which takes $O(\log^2 p + \log q)$ steps per operation,
and Attiya and Fouren's $\Omega(\min(c,\log\log p))$ lower bound \cite{DBLP:conf/opodis/AttiyaF17}.
It would be interesting to determine how the true step complexity of lock-free queues (or, more generally, bags)
depends on $p$.
Since a queue is also a bag, our queue is the first lock-free bag we know of that has polylogarithmic step complexity.

We believe the approach used here to implement a lock-free queue 
could be applied to obtain other lock-free
data structures with a polylogarithmic step complexity.
For example, we can easily adapt our routines to implement a  vector data structure that stores a sequence and
provides three operations: \opa{Append}{e} to add an element \var{e} to the end of the sequence,
\opa{Get}{i} to read the \var{i}th element in the sequence, and
\opa{Index}{e} to compute the position of element \var{e} in the sequence.
%\here{Can omit rest of this paragraph to save space}
%An \opa{append}{e} is implemented like \opa{Enqueue}{e} in $O(\log p)$ steps.  
%A \opa{get}{i} is similar to \op{GetEnqueue}, taking $O(\log n + \log^2p)$ steps when the vector has $n$ elements.  
%An \opa{index}{e} is similar to \op{IndexDequeue} (except operating on enqueues instead of dequeues) and would take $O(\log p)$ steps.
\here{doublecheck all preceding claims for accuracy.}
We are investigating whether a similar approach can be used for stacks, deques or even priority queues.

>>>>

% !TEX root =  queue.tex

@@

It would be interesting to close the gap that remains between our queue, which takes $O(\log^2 p + \log q)$ steps per operation,
and Jayanti, Tarjan and Boix-Adser\`{a}'s $\Omega(\log p)$ lower bound \cite{JTB19}.
For the more relaxed bag data structure the gap is larger between the  $\Omega(\min(c,\log\log p))$ lower bound \cite{DBLP:conf/opodis/AttiyaF17} and our upper bound of $O(\log^2 p + \log q)$.
Could the complexity for either queues or bags be made polylogarithmic in $p$ while being independent of the size $q$ of the data structure?

We believe the approach used here to implement a lock-free queue 
could be applied to obtain other lock-free
data structures with a polylogarithmic step complexity.
For example, we can easily adapt our routines to implement a  vector data structure that stores a sequence and
provides three operations: \opa{Append}{e} to add an element \var{e} to the end of the sequence,
\opa{Get}{i} to read the \var{i}th element in the sequence, and
\opa{Index}{e} to compute the position of element \var{e} in the sequence.
%\here{Can omit rest of this paragraph to save space}
%An \opa{append}{e} is implemented like \opa{Enqueue}{e} in $O(\log p)$ steps.  
%A \opa{get}{i} is similar to \op{GetEnqueue}, taking $O(\log n + \log^2p)$ steps when the vector has $n$ elements.  
%An \opa{index}{e} is similar to \op{IndexDequeue} (except operating on enqueues instead of dequeues) and would take $O(\log p)$ steps.
\here{doublecheck all preceding claims for accuracy.}
Building on the work described in this paper, Asbell and Ruppert \cite{AR23} have
designed stacks and doubly-ended queues with polylogarithmic amortized step
complexity.  This required a substantially different representation of the data stored in the ordering tree.
\here{?  They also showed that a randomized implementation of the ordering tree technique can
be used to obtain a (doubly-ended) queue with O(\here{?}) amortized expected step complexity,
beating the  $\Omega(\min(c,\log\log p))$ lower bound on deterministic algorithms.}
Whether the ordering tree could also be used to obtain a priority queue with polylogarithmic step complexity
remains an open question.

### 2023-06-29 journal/related.tex
Updates to related work--incorporating new info about Jayanti et al.'s lower bound

>>>>

\paragraph{Lower Bounds.}
As mentioned in the introduction, it follows from Attiya and Fouren's lower bound on bag data structures
\cite{DBLP:conf/opodis/AttiyaF17} that the
amortized step complexity of operations on a queue is $\Omega(\min(c,\log\log p))$, where $c$ is contention.
Subsequently, Jayanti, Tarjan and Boix-Adser\`{a} \cite{JTB19} proved a stronger $\Omega(\log p)$ lower bound on
the amortized step complexity of queue operations.
